Equalizer\\Programming and User Guide

### 2010-02-02 Developer/ProgrammingGuide/paper.tex
Programming Guide: fix typo
[ ] May break build
[ ] May break existing applications
[x] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[ ] Documentation
Update Programming Guide and 0.9.1-related web pages
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.2 for Equalizer 0.9+svn\\[\medskipamount]
%  \today
}

@@

\lstset{language=C++}
\section*{Equalizer 0.9 Programming and User Guide}
\today
\subsection*{Contributors}

@@

\textcopyright 2007-2009 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

@@

The images on the back page show the following scalable rendering modes:
Combined 2D/DB [top left], DPlex [middle left], Pixel [middle center],
2D [bottom left], DB [bottom center] and stereo [bottom right].

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures \vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    \hline
  \end{tabularx}}
\clearpage

@@

For each node, create on \textsf{pipe} for each graphics card in the
machine. Set the \textsf{device} number to the correct index. On
operating systems using X11, e.g., Linux, also set the port number if
your X-Server is running on a nonstandard port.

@@

\begin{figure}[h!t]
  \subfigure[]{\label{fMono}
    \includegraphics[height=2.5cm]{images/mono}}\hfil
  \subfigure[]{\label{fStereo}
    \includegraphics[height=2.5cm]{images/stereo}}\hfil
  \subfigure[]{\label{fTracked}
    \includegraphics[height=2.5cm]{images/tracked}}
  \subfigure[]{\label{fHMD}
    \includegraphics[height=2.5cm]{images/hmd}}%
  {\caption{\label{fImmersive}Monoscopic, Stereoscopic, Tracked and HMD
    Frusta}}
\end{figure}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
  \vspace{-3ex}
\end{wrapfigure}
To provide a motivation for the design of the image compositing
API, the direct send parallel compositing algorithm is introduced in this
section. Other parallel compositing algorithms, e.g. binary-swap, can
also be expressed through an Equalizer configuration file.

The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
The frame data is a holder for images and additional information, such
as output frame attributes and pixel data availability.

@@

The frame synchronization through the frame data allows the input frame
to wait for the pixel data to become ready, which is signaled by the
output frame after readback.

@@

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/compositor.pdf}
  {\caption{\label{fCompositor}Functional Diagram of the Compositor}}
\end{wrapfigure}
On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semitransparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the pre-integrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

@@

The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{center}
\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_HOSTNAME\\
  \textbf{Value} & string\\
  \textbf{Default} & ''localhost''\\
  \textbf{Details} & The hostname or IP address used to connect the
  server or node. When used for the server, the listening port of the
  server is bound to this address. When used for a node, the server
  first tries to connect to the render client node using this hostname,
  and then tries to launch the render client executable on this host.\\
  \textbf{See also} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND
  EQ\_CONNECTION\_IATTR\_TCPIP\_PORT EQ\_CONNECTION\_IATTR\_TYPE\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Details} & The command used by the server to auto-launch nodes
  which could not be connected. The launch command is executed from a
  process forked from the server process. The \% tokens are replaced by
  the server at runtime with concrete data: \%h is replaced by the
  hostname, \%c by the render client command to launch, including command line
  arguments and \%n by a node-unique identifier. Each command line
  argument is surrounded by launch command quotes.\\
  \textbf{See also} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND\_QUOTE 
  EQ\_CONNECTION\_IATTR\_LAUNCH\_TIMEOUT\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Details} & The server uses command line arguments to launch
  render client nodes correctly. Certain launch commands or shells use
  different conventions to separate command line arguments. These
  arguments might contain white spaces, and therefore have to be
  surrounded by quotes to identify their borders. This option is mostly
  used on Windows.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP \textbar \ PIPE [Win32 only]\\
  \textbf{Default} & TCPIP\\
  \textbf{Details} & The protocol for connections. SDP programmatically
  selects the socket direct protocol (AF\_INET\_SDP) provided by most
  InfiniBand protocol stacks, TCPIP uses normal TCP
  sockets (AF\_INET). PIPE uses a named pipe to communicate between
  two processes on the same machine.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TCPIP\_PORT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 0\\
  \textbf{Details} & The listening port used by the server or
  node. For nodes, the port can be used to contact pre-started, resident
  render client nodes or to use a specific port for the node. If 0 is
  specified, a random port is chosen. Note that a server with no
  connections automatically creates a default connection using the
  server's default port.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_PIPE\_FILENAME\\
  \textbf{Value} & string\\
  \textbf{Default} & none\\
  \textbf{Details} & The filename of the named pipe used by the server or
  node. The filename has to be unique on the local host.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_LAUNCH\_TIMEOUT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 60'000 (1 minute)\\
  \textbf{Details} & Defines the timeout in milliseconds to wait for
  an auto-launched node. If the render client process did not contact
  the server within that time, the node is considered to be unreachable
  and the initialization of the configuration fails.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & See EQ\_WINDOW\_IATTR\_HINT\_STATISTICS.\\
  \textbf{See also} & EQ\_NODE\_IATTR\_HINT\_STATISTICS 
  EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \hline
\end{tabularx}\\\vfill

@@

A connection description defines the network parameters of an Equalizer
process. Currently TCP/IP, SDP and PIPE connection types are
supported. TCP/IP creates a TCP socket. SDP is very similar, except that
the address family \textsf{AF\_INET\_SDP} instead of \textsf{AF\_INET}
is used to enforce a SDP connection. PIPE uses a named pipe for fast
interprocess communication on Windows.

@@

The hostname is the IP address or (resolvable) host name. A server or
node may have multiple connection descriptions, for example to use a
named pipe for local communications and TCP/IP for remote nodes.

@@

{\footnotesize\begin{lstlisting}
server
{
    connection # 0-n times, listening connections of the server
    {
        type       TCPIP | SDP | PIPE [Win32 only]
        TCPIP_port unsigned
        PIPE_filename string
        hostname   string
    }

@@

    load_equalizer # adjust 2D tiling or DB range of children
    {
        mode 2D | DB | VERTICAL | HORIZONTAL
        damping float # 0: no damping, 1: no changes
    }
    DFR_equalizer # adjust ZOOM to achieve constant framerate
    {
        framerate float # target framerate
        damping float   # 0: no damping, 1: no changes
    }
    framerate_equalizer {} # smoothen window swapbuffer rate (DPlex)
    monitor_equalizer {}   # set frame zoom when monitoring other views

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.3 for Equalizer 0.9.1\\[\medskipamount]
%  \today
}

@@

\lstset{language=C++}
\section*{Equalizer 0.9.1 Programming and User Guide}
\today
\subsection*{Contributors}

@@

\textcopyright 2007-2010 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

@@

The images on the back page show the following scalable rendering modes:
Subpixel (FSAA) [top left], DPlex [middle left], Pixel [middle center],
2D [bottom left], DB [bottom center] and stereo [bottom right].

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    \hline
  \end{tabularx}}
\clearpage

@@

For each node, create one \textsf{pipe} for each graphics card in the
machine. Set the \textsf{device} number to the correct index. On
operating systems using X11, e.g., Linux, also set the port number if
your X-Server is running on a nonstandard port.

@@

\begin{figure}[h!t]\center
  \subfigure[]{\label{fMono}
    \includegraphics[height=4cm]{images/mono}}\hfil
  \subfigure[]{\label{fStereo}
    \includegraphics[height=4cm]{images/stereo}}
  \subfigure[]{\label{fTracked}
    \includegraphics[height=4cm]{images/tracked}}\hfil
  \subfigure[]{\label{fHMD}
    \includegraphics[height=4cm]{images/hmd}}
  {\caption{\label{fImmersive}Monoscopic (a), Stereoscopic (b), Tracked (c) and
      HMD (d) Frusta}}
\end{figure}

@@

To provide a motivation for the design of the image compositing
API, the direct send parallel compositing algorithm is introduced in this
section. Other parallel compositing algorithms, e.g. binary-swap, can
also be expressed through an Equalizer configuration file.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

@@

The frame data is a holder for images and additional information, such
as output frame attributes and pixel data availability.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
The frame synchronization through the frame data allows the input frame
to wait for the pixel data to become ready, which is signaled by the
output frame after readback.

@@

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/compositor.pdf}
  {\caption{\label{fCompositor}Functional Diagram of the Compositor}}
\end{figure}
On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semitransparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the pre-integrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{center}
\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_HOSTNAME\\
  \textbf{Value} & string\\
  \textbf{Default} & ''localhost''\\
  \textbf{Details} & The hostname or IP address used to connect the
  server or node. When used for the server, the listening port of the
  server is bound to this address. When used for a node, the server
  first tries to connect to the render client node using this hostname,
  and then tries to launch the render client executable on this host.\\
  \textbf{See also} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND
  EQ\_CONNECTION\_IATTR\_PORT
  EQ\_CONNECTION\_IATTR\_TYPE\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP \textbar \ RSP \textbar \ 
  PIPE [Win32 only]\\
  \textbf{Default} & TCPIP\\
  \textbf{Details} & The protocol for connections. SDP programmatically
  selects the socket direct protocol (AF\_INET\_SDP) provided by most
  InfiniBand protocol stacks, TCPIP uses normal TCP
  sockets (AF\_INET). RSP provides reliable UDP multicast. PIPE uses a named
  pipe to communicate between two processes on the same machine.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_PORT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 0\\
  \textbf{Details} & The listening port used by the server or
  node. For nodes, the port can be used to contact pre-started, resident
  render client nodes or to use a specific port for the node. If 0 is
  specified, a random port is chosen. Note that a server with no
  connections automatically creates a default connection using the
  server's default port.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_FILENAME\\
  \textbf{Value} & string\\
  \textbf{Default} & none\\
  \textbf{Details} & The filename of the named pipe used by the server or
  node. The filename has to be unique on the local host.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Details} & The command used by the server to auto-launch nodes
  which could not be connected. The launch command is executed from a
  process forked from the server process. The \% tokens are replaced by
  the server at runtime with concrete data: \%h is replaced by the
  hostname, \%c by the render client command to launch, including command line
  arguments and \%n by a node-unique identifier. Each command line
  argument is surrounded by launch command quotes.\\
  \textbf{See also} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND\_QUOTE 
  EQ\_NODE\_IATTR\_LAUNCH\_TIMEOUT\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Details} & The server uses command line arguments to launch
  render client nodes correctly. Certain launch commands or shells use
  different conventions to separate command line arguments. These
  arguments might contain white spaces, and therefore have to be
  surrounded by quotes to identify their borders. This option is mostly
  used on Windows.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_IATTR\_LAUNCH\_TIMEOUT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 60'000 (1 minute)\\
  \textbf{Details} & Defines the timeout in milliseconds to wait for
  an auto-launched node. If the render client process did not contact
  the server within that time, the node is considered to be unreachable
  and the initialization of the configuration fails.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & See EQ\_WINDOW\_IATTR\_HINT\_STATISTICS.\\
  \textbf{See also} & EQ\_NODE\_IATTR\_HINT\_STATISTICS 
  EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \hline
\end{tabularx}\\%\vfill

@@

A connection description defines the network parameters of an Equalizer
process. Currently TCP/IP, SDP, RSP and PIPE connection types are
supported. TCP/IP creates a TCP socket. SDP is very similar, except that the
address family \textsf{AF\_INET\_SDP} instead of \textsf{AF\_INET} is used to
enforce a SDP connection. RSP provides reliable UDP multicast. PIPE uses a named
pipe for fast interprocess communication on Windows.

@@

The hostname is the IP address or resolvable host name. A server or
node may have multiple connection descriptions, for example to use a
named pipe for local communications and TCP/IP for remote nodes.

The interface is the IP address or resolvable host name of the adapter to which
multicast traffic is send.

@@

{\footnotesize\begin{lstlisting}
server
{
    connection # 0-n times, listening connections of the server
    {
        type       TCPIP | SDP | RSP | PIPE [Win32 only]
        port       unsigned
        filename   string
        hostname   string
        interface  string
    }

@@

    load_equalizer # adjust 2D tiling or DB range of children
    {
        mode 2D | DB | VERTICAL | HORIZONTAL
        damping float # 0: no damping, 1: no changes
        boundary [ x y ] | float  # x,y: tile boundary for 2D, float: range
    }
    DFR_equalizer # adjust ZOOM to achieve constant framerate
    {
        framerate float # target framerate
        damping float   # 0: no damping, 1: no changes
    }
    framerate_equalizer {} # smoothen window swapbuffer rate (DPlex)
    monitor_equalizer {}   # set frame zoom when monitoring other views
